sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-11-23 doc/paper.Rmd
Try adding spBayes

<!-- TODO: cite:  -->
<!-- @article{gelfand2017, -->
<!--   title = {Bayesian Modeling and Analysis of Geostatistical Data}, -->
<!--   author = {Gelfand, Alan E. and Banerjee, Sudipto}, -->
<!--   year = {2017}, -->
<!--   journal = {Annual Review of Statistics and Its Application}, -->
<!--   volume = {4}, -->
<!--   number = {1}, -->
<!--   pages = {245--266}, -->
<!--   doi = {10.1146/annurev-statistics-060116-054155}, -->
<!--   annotation = {\_eprint: https://doi.org/10.1146/annurev-statistics-060116-054155} -->
<!-- } -->

@@

sdmTMB fits GLMMs to spatial or spatiotemporal geostatistical data.
Geostatistical data simply refers to data observed at specific spatial coordinates (vs. areal data that might represent polygons or grids).
The process component of an sdmTMB model can be formed by any combination of main ('fixed') effects, spatial random fields, spatiotemporal random fields, IID random intercepts, time-varying effects, and spatially varying effects (Fig. \@ref(fig:diagram)).
This process component is combined with an observation error family (e.g., Gaussian, Gamma, binomial, Tweedie) and link (e.g., identity, log, logit).

@@

```{r compare-table}
library(kableExtra)
suppressMessages(suppressWarnings(d <- readr::read_csv(here::here("doc/software-comparison.csv"), show_col_types = FALSE, comment = "#")))
d <- as.matrix(d)
d[d == "1"] <- "\\checkmark"
d[d == "1*"] <- "\\checkmark\\footnote{a}"
d[d == "0"] <- "--"
d[d == "0*"] <- "--\\footnote{a}"
d <- as.data.frame(d)
names(d)[1] <- ""
row.names(d) <- NULL
d <- d[,c(1, 2, 3, 5, 6, 4)]
d <- d[!grepl("SPDE", d[,1]), ]
names(d)[names(d) == "mgcv"] <- "mgcv\\footnote{a}"

knitr::kable(d, format = "latex", caption = "Comparison of functionality between five R packages that can fit SPDE-based geostatistical GLMMs.", 
  booktabs = TRUE, 
  linesep = c(
    rep('', 4), # coefs
    '\\addlinespace',
    rep('', 3), # correlation stuff
    '\\addlinespace',
    rep('', 10), # likelihood
    '\\addlinespace',
    rep('', 2), # bayes
    '\\addlinespace'
  ),
  escape = FALSE, row.names = FALSE) %>%
  kableExtra::kable_styling(font_size = 8) %>% 
  kableExtra::footnote(
    number = c(
      "SPDE approach as in Miller et al. (2019).",
      "Technically possible but non-trivial.",
      "Zero-inflated NB2 only.",
      "Tweedie power parameter fixed for \\\\texttt{mgcv::gamm()}.",
      "Technically possible but non-trivial.",
      "Technically possible but non-trivial."
    ),
    escape = FALSE
  )
```

>>>>

<!-- TODO: cite:  -->
<!-- @article{gelfand2017, -->
<!--   title = {Bayesian Modeling and Analysis of Geostatistical Data}, -->
<!--   author = {Gelfand, Alan E. and Banerjee, Sudipto}, -->
<!--   year = {2017}, -->
<!--   journal = {Annual Review of Statistics and Its Application}, -->
<!--   volume = {4}, -->
<!--   number = {1}, -->
<!--   pages = {245--266}, -->
<!--   doi = {10.1146/annurev-statistics-060116-054155}, -->
<!--   annotation = {\_eprint: https://doi.org/10.1146/annurev-statistics-060116-054155} -->

<!-- @book{diggle2007, -->
<!--   title = {Model-Based {{Geostatistics}}}, -->
<!--   author = {Diggle, Peter and Ribeiro, Paulo Justiniano}, -->
<!--   year = {2007}, -->
<!--   publisher = {{Springer}}, -->
<!--   googlebooks = {qCqOm39OuFUC}, -->
<!--   isbn = {978-0-387-48536-2}, -->
<!--   langid = {english} -->
<!-- } -->

<!-- } -->

@@

sdmTMB fits GLMMs to spatial or spatiotemporal geostatistical data.
Geostatistical data simply refers to data observed at specific spatial coordinates (vs. areal data that might represent polygons or grids).
The data can be collected across discrete points in time.
The process component of an sdmTMB model can be formed by any combination of main ('fixed') effects, spatial random fields, spatiotemporal random fields, IID random intercepts, time-varying effects, and spatially varying effects (Fig. \@ref(fig:diagram)).
This process component is combined with an observation error family (e.g., Gaussian, Gamma, binomial, Tweedie) and link (e.g., identity, log, logit).

@@

```{r compare-table}
library(kableExtra)
suppressMessages(suppressWarnings(d <- readr::read_csv(here::here("doc/software-comparison.csv"), show_col_types = FALSE, comment = "#")))
d <- as.matrix(d)
d[d == "1"] <- "\\checkmark"
d[d == "1*"] <- "\\checkmark\\footnote{a}"
d[d == "0"] <- "--"
d[d == "0*"] <- "--\\footnote{a}"
d <- as.data.frame(d)
names(d)[1] <- ""
row.names(d) <- NULL
d <- d[,c(1, 2, 3, 5, 6, 4, 7)]
d <- d[!grepl("SPDE", d[,1]), ]
names(d)[names(d) == "mgcv"] <- "mgcv\\footnote{a}"

knitr::kable(d, format = "latex", caption = "Comparison of functionality between six R packages that can fit geostatistical GLMMs.", 
  booktabs = TRUE, 
  linesep = c(
    rep('', 4), # coefs
    '\\addlinespace',
    rep('', 4), # correlation stuff
    '\\addlinespace',
    rep('', 10), # likelihood
    '\\addlinespace',
    rep('', 2), # bayes
    '\\addlinespace'
  ),
  escape = FALSE, row.names = FALSE) %>%
  kableExtra::kable_styling(font_size = 8) %>% 
  kableExtra::footnote(
    number = c(
      "SPDE approach as in Miller et al. (2019).",
      "Technically possible but non-trivial.",
      "Zero-inflated NB2 only.",
      "Tweedie power parameter fixed for \\\\texttt{mgcv::gamm()}.",
      "Technically possible but non-trivial.",
      "Technically possible but non-trivial."
    ),
    escape = FALSE
  )
```
